{\steve}: A hierarchical Bayesian model for Supernova Cosmology

### 2018-10-15 papers/methods/paper.tex
emulateapj tables and abstract comments

\usepackage{lineno}
\linenumbers

\graphicspath{ {./figures/} }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

 % If you need two or more lines of authors, add an extra line using \newauthor
 \author{
 	S.~R.~Hinton\altaffilmark{1,2},\thanks{E-mail: samuelreay@gmail.com}
 	T.~M.~Davis\altaffilmark{1,2},
 	A.~G.~Kim\altaffilmark{3},
 	A.~M\"oller\altaffilmark{2,4},
 	M.~Sako\altaffilmark{5},
 	M.~Smith\altaffilmark{6},
 	\\}
 \affil{$^{1}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
 \affil{$^{2}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
 \affil{$^{3}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
 \affil{$^{4}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
 \affil{$^{5}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
 \affil{$^{6}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}

% Abstract of the paper
\begin{abstract}
We present a new Bayesian hierarchical model named {\steve} for performing type Ia supernova (SN~Ia) cosmology fits. 
%
This advances previous works with Bayesian hierarchical models by including 
an improved treatment of Malmquist bias, 
accounting for additional sources of systematic uncertainty, 
and increasing numerical efficiency. 
%
Given light curve fit parameters, redshifts, and host-galaxy masses, we fit {\steve} simultaneously for parameters describing cosmology, SN~Ia populations, and systematic uncertainties. Selection effects are characterised using Monte-Carlo simulations.
%
We demonstrate its implementation by fitting realisations of SN~Ia datasets where the SN~Ia model closely follows that used in {\steve}.
%
Next, we validate on more realistic SNANA simulations of SN~Ia samples from the Dark Energy Survey and low-redshift surveys \citep{DESKEY}.
%
These simulated datasets contain more than $60\,000$ SNe~Ia, which we use to evaluate biases in the recovery of cosmological parameters, specifically the equation-of-state of dark energy, $w$. 
%
This is the most rigorous test of a BHM model, and reveals small $w$-biases that depend on the simulated SN~Ia properties, in particular the intrinsic SN~Ia scatter model. This $w$-bias is less than $0.03$ on average, less than half the statistical uncertainty on $w$.
%
These simulation test results are a concern for BHM applications on large upcoming surveys, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
\end{abstract}

@@

\begin{eqnarray}
\mu_C &= 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
d_L& = (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
E(z) &= \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
\end{eqnarray}

@@

\begin{table}
	\centering
	\caption{ Parameters defined in our model and a summary of their use. The parameters are broken into multiple sections, the top for parameters which are defined globally across all surveys, the second section for parameters which are defined for each survey, the third section are parameters which are defined for each supernova. The bottom section shows the observables (not parameters) that are the input data to the model.}
	\label{tab:param_summary}
	\begin{tabular}{ll}
		\hline
		Parameter & Description \\
		\hline
		$\Omega_m$  & Matter density  \\
		$w$  &  Dark energy equation of state  \\
		$\alpha$ & Stretch standardisation   \\
		$\beta$         &  color standardisation   \\
		$\delta(0)$ & Scale of the mass-magnitude correction\\
		$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
		$\delta\mathcal{Z}_i$ & Systematics scale\\
		$\langle M_B \rangle$ & Mean absolute magnitude \\
		%$s_c$ & color bias\\
		\hline
		$\delta S$ & Selection effect deviation \\
		$\langle x_1^i \rangle$ & Mean stretch nodes\\
		$\langle c^i \rangle$ & Mean color nodes\\
		$\alpha_c$ & Skewness of color population \\
		$\sigma_{M_B}$ & Population magnitude scatter \\
		$\sigma_{x_1}$ & Population stretch scatter \\
		$\sigma_{c}$ & Population color scatter \\
		$\kappa_{0}$ & Extra color dispersion\\
		$\kappa_{1}$ & Redshift-dependence of extra color dispersion\\
		\hline
		$m_B$ & True flux\\
		$x_1$ & True stretch \\
		$c$ & True color \\
		$z$ & True redshift \\
		$M_B$ & Derived absolute magnitude \\
		$\mu$ & Derived distance modulus \\
		\hline
		$\hat{m}_B$ & Measured flux\\
		$\hat{x}_1$ & Measured stretch \\
		$\hat{c}$ & Measured color \\
		$C$ & Covariance on flux, stretch and color \\
		$\hat{z}$ & Observed redshift \\
		$\hat{m}$ & Determined mass probability \\
		$\hat{p}$ & Determined Ia probability \\
		\hline
	\end{tabular}
\end{table}

@@

\begin{eqnarray}
m_B^*(z) &= \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
\sigma^{*2}_{m_B} &= \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
\end{eqnarray}

@@

\begin{eqnarray}
d_{\rm CDF} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
\end{eqnarray}

@@

\begin{table}
	\centering
	\caption{Cosmological parameters determined from the surfaces of 100 fits to independent realisations of toy supernova data. As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae. For each chain, we record the mean and standard deviation, and then show the average mean and average standard deviation in the table. The scatter introduced by simulation variance (the standard deviation of the 100 mean parameter values) is shown in brackets. Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$. As we are using 100 independent realisations, the precision of our determination of the mean simulation result is a tenth of the quoted standard deviation: $\sqrt{100} = 10$. As the deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model. For the Flat $w$CDM model, the value of $w$ is reported with a prior on $\Omega_m$ of $\mathcal{N}(0.3, 0.01)$.}
	\label{tab:simple_model}
	\begin{tabular}{l|cc}
		\hline
		Model & $\Omega_m\, \langle\mu\rangle,\langle\sigma\rangle$(scatter) & $w\, \langle\mu\rangle,\langle\sigma\rangle$(scatter) \\ 
		\hline
		Flat $\Lambda$CDM & $0.301,\, 0.015\,(0.012)$ & -- \\ 
		Flat $w$CDM & -- & $-1.00,\, 0.042\,(0.030)$ \\ 
		\hline
	\end{tabular}
\end{table}

@@

To investigate biases in the model in fine detail, we look for systematic bias in $\Omega_m$ in the Flat $\Lambda$CDM cosmology test, and bias in $w$ for the Flat $w$CDM test with strong prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$. This allows us to investigate biases without the investigative hindrances of non-Gaussian or truncated posterior surfaces. The strong prior on $\Omega_m$ cuts a slice through the traditional `banana' posterior surface in the $w$-$\Omega_m$ plane of Figure \ref{fig:simple_w}. Without making such a slice, the variation in $w$ is larger due to a shift along the degeneracy direction of the `banana'. By focusing the slice at an almost fixed $\Omega_m$, we can see the variation in the mean value of $w$ approximately perpendicular to the lines of degeneracy, instead of along them. The results of the analysis are detailed in Table \ref{tab:simple_model}, and demonstrate the performance of our model in recovering the true cosmological parameters. With this simple data, we also correctly recover underlying supernova populations, which can be seen in Figure \ref{fig:simple_w_super}.

@@

\begin{table}
	\centering
	\caption{Tested population distributions, where the {\sk} low-$z$ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
	\label{tab:dist}
	\begin{tabular}{l|cccc}
		\hline
		Model & $\langle x_1 \rangle$ & $\sigma_{x_1}$ &  $\langle c \rangle$ & $\sigma_c$  \\
		\hline
		{\sk} low-$z$    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\
		{\sk} DES     & $0.973$ & $^{+0.222}_{1.472}$ & $-0.054$  &  $^{+0.101}_{0.043}$ \\
		\hline
	\end{tabular}
\end{table}

@@

\begin{table*}
	\centering
	\caption{Standardisation parameters and base intrinsic scatter parameter results for the 100 fits to {\gten} and {\celeven} simulations. We show the average parameter mean and average standard deviation respectively, with the simulation scatter shown in brackets, such that each cell shows $\langle\mu\rangle\ [\langle\sigma\rangle\ ({\rm scatter})]$. The width of the intrinsic scatter ($\sigma_{m_B}$) does not have an input truth value as it is determined from the scatter model.}
	\label{tab:standardisation}
	\begin{tabular}{l|rrrrr}
		\hline
		Model & $\alpha - \alpha_{\rm True}$ & $\beta - \beta_{\rm True}$ & $\langle M_B \rangle - \langle M_B \rangle_{\rm True}$ & $\sigma_{m_B}^{\rm DES}$ & $\sigma_{m_B}^{{\rm low-}z}$ \\
		\hline
		{\gten} Stat + Syst  & $0.022 \ [0.009\ (0.008)]$ & $ 0.34 \ [0.19\ (0.18)]$ & $-0.002 \ [0.028\ (0.015)]$ & $0.070 \ [0.022\ (0.018)]$ & $0.073 \ [0.025\ (0.022)]$\\
		{\gten}  Stat         & $0.000 \ [0.008\ (0.008)]$ & $0.33 \ [0.20\ (0.17)]$ & $0.001 \ [0.016\ (0.013)]$ & $0.069 \ [0.023\ (0.019)]$ & $0.072 \ [0.026\ (0.023)]$\\
		{\celeven} Stat + Syst  &  $0.002 \ [0.009\ (0.007)]$ & $-0.04 \ [0.15\ (0.13)]$ & $0.014 \ [0.030\ (0.018)]$ & $0.024 \ [0.016\ (0.011)]$ & $0.029 \ [0.020\ (0.014)]$ \\
		{\celeven} Stat         &  $0.000 \ [0.008\ (0.007)]$ & $-0.05 \ [0.16\ (0.13)]$ & $0.006 \ [0.016\ (0.015)]$ & $0.025 \ [0.016\ (0.012)]$ & $0.027 \ [0.020\ (0.015)]$\\
		\hline
	\end{tabular}
\end{table*}

@@

\begin{table}
	\centering
	\caption{Investigating the combined 100 fits to {\gten} and {\celeven} simulations, fitting with both statistics only and also when including systematics. The quoted value for $w$ represents the average mean of the fits, with the average uncertainty being shown in square brackets and the simulation scatter (the standard deviation of the mean of 100 fits) shown in standard brackets. The bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation.}
	\label{tab:bulk_summary}
	\begin{tabular}{l|rr}
		\hline
		Model & $w\ \langle\mu\rangle\ [\langle\sigma_w\rangle$ (scatter)$]$ & $w$-Bias \\
		\hline
		{\gten} Stat + Syst  &  $-0.998 \ [0.097\ (0.073)]$ & $(0.02\pm0.07)\sigma$  \\
		{\gten}  Stat         &  $-1.008\ [0.080\ (0.068)]$ & $(-0.10\pm0.08)\sigma$   \\
		{\celeven} Stat + Syst  &  $-0.945\ [0.098\ (0.077)]$ & $(0.55\pm0.08)\sigma$  \\
		{\celeven} Stat         &  $-0.948\ [0.079\ (0.066)]$ & $(0.65\pm0.08)\sigma$   \\
		\hline
	\end{tabular}
\end{table}

>>>>

\usepackage{lineno}
\linenumbers
\renewcommand{\arraystretch}{1.2}
\graphicspath{ {./figures/} }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\author{
	S.~R.~Hinton\altaffilmark{1},
	T.~M.~Davis\altaffilmark{1},
	A.~G.~Kim\altaffilmark{3},
	R.~Kessler\altaffilmark{2},
	C.~Lidman\altaffilmark{4},
	A.~M\"oller\altaffilmark{5,4},
	M.~Sako\altaffilmark{6},
	D.~Scolnic\altaffilmark{2},
	M.~Smith\altaffilmark{7} 
	\\ \vspace{0.2cm} (DES Collaboration) \\	
}

\affil{$^{1}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia}
\affil{$^{2}$ Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}
\affil{$^{3}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}
\affil{$^{4}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia}
\affil{$^{5}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
\affil{$^{6}$ Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, PA 19104, USA}
\affil{$^{7}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}

% Abstract of the paper
\begin{abstract}
We present a new Bayesian hierarchical model (BHM) named {\steve} for performing type Ia supernova (SN~Ia) cosmology fits. 
%
This advances previous works by including 
an improved treatment of Malmquist bias, 
accounting for additional sources of systematic uncertainty, 
and increasing numerical efficiency. 
%
Given light curve fit parameters, redshifts, and host-galaxy masses, we fit {\steve} simultaneously for parameters describing cosmology, SN~Ia populations, and systematic uncertainties. Selection effects are characterised using Monte-Carlo simulations.
%
We demonstrate its implementation by fitting realisations of SN~Ia datasets where the SN~Ia model closely follows that used in {\steve}.
%
Next, we validate on more realistic SNANA simulations of SN~Ia samples from the Dark Energy Survey and low-redshift surveys \citep{DESKEY}.
%
These simulated datasets contain more than $60\,000$ SNe~Ia, which we use to evaluate biases in the recovery of cosmological parameters, specifically the equation-of-state of dark energy, $w$. 
%
This is the most rigorous test of a BHM method applied to SN~Ia cosmology fitting, and reveals small $w$-biases that depend on the simulated SN~Ia properties, in particular the intrinsic SN~Ia scatter model. This $w$-bias is less than $0.03$ on average, less than half the statistical uncertainty on $w$.
%
These simulation test results are a concern for BHM cosmology fitting applications on large upcoming surveys, and therefore future development will focus on minimising the sensitivity of {\steve} to the SN~Ia intrinsic scatter model.
\end{abstract}

@@

\begin{eqnarray}
\mu_C &=& 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
d_L &=& (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
E(z) &=& \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
\end{eqnarray}

@@

\begin{deluxetable}{ll}
	\tablecolumns{2}
	\tablewidth{0.99\columnwidth}
	\tablecaption{Model Parameters}
	\tablehead { \colhead{Parameter} & \colhead{Description} }
	\startdata

	&\textbf{Global Parameters} \\

	$\Omega_m$  & Matter density  \\
	$w$  &  Dark energy equation of state  \\
	$\alpha$ & Stretch standardisation   \\
	$\beta$         &  color standardisation   \\
	$\delta(0)$ & Scale of the mass-magnitude correction\\
	$\delta(\infty)/\delta(0)$ & Redshift-dependence of mass-magnitude correction\\
	$\delta\mathcal{Z}_i$ & Systematics scale\\
	$\langle M_B \rangle$ & Mean absolute magnitude \\
	\\
	& \textbf{Survey Parameters} \\

	$\delta S$ & Selection effect deviation \\
	$\langle x_1^i \rangle$ & Mean stretch nodes\\
	$\langle c^i \rangle$ & Mean color nodes\\
	$\alpha_c$ & Skewness of color population \\
	$\sigma_{M_B}$ & Population magnitude scatter \\
	$\sigma_{x_1}$ & Population stretch scatter \\
	$\sigma_{c}$ & Population color scatter \\
	$\kappa_{0}$ & Extra color dispersion\\
	$\kappa_{1}$ & Redshift-dependence of extra color dispersion\\
	\\
	& \textbf{Supernova Parameters} \\
	$m_B$ & True flux\\
	$x_1$ & True stretch \\
	$c$ & True color \\
	$z$ & True redshift \\
	$M_B$ & Derived absolute magnitude \\
	$\mu$ & Derived distance modulus \\
	\\
	& \textbf{Input Data}\textsuperscript{a} \\

	$\hat{m}_B$ & Measured flux\\
	$\hat{x}_1$ & Measured stretch \\
	$\hat{c}$ & Measured color \\
	$C$ & Covariance on flux, stretch and color \\
	$\hat{z}$ & Observed redshift \\
	$\hat{m}$ & Determined mass probability \\
	$\hat{p}$ & Determined Ia probability \\
	\enddata
	\tablenotetext{a}{Not model parameters but shown for completeness.}
\label{tab:param_summary}
\end{deluxetable}

@@

\begin{eqnarray}
m_B^*(z) &=& \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
\sigma^{*2}_{m_B} &=& \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
\end{eqnarray}

@@

\begin{eqnarray}
d_{\rm CDF} &=& \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &=& 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad& \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
\end{eqnarray}

@@

\begin{deluxetable}{lcc}
	\tablecolumns{3}
	\tablecaption{Cosmological Parameters for toy supernova data}
	\tablehead {
		\colhead {Model}              &
		\colhead {$\Omega_m\, \langle\mu\rangle,\langle\sigma\rangle$(scatter)}  &
		\colhead{$w\, \langle\mu\rangle,\langle\sigma\rangle$(scatter)}
	}

	\startdata
		Flat $\Lambda$CDM & $0.301,\, 0.015\,(0.012)$ & -- \\ 
		Flat $w$CDM & -- & $-1.00,\, 0.042\,(0.030)$ \\ 
	\enddata
	\label{tab:simple_model}
	\tablecomments{Cosmological parameters determined from the surfaces of 100 fits to independent realisations of toy supernova data. As described in the main text, each dataset comprised 1000 low-redshift supernovae and 1000 high-redshift supernovae. For each chain, we record the mean and standard deviation, and then show the average mean and average standard deviation in the table. The scatter introduced by simulation variance (the standard deviation of the 100 mean parameter values) is shown in brackets. Model bias would appear as shifts away from the simulation values of $\Omega_m = 0.3$, $w = -1$.}
\end{deluxetable}

@@

To investigate biases in the model in fine detail, we look for systematic bias in $\Omega_m$ in the Flat $\Lambda$CDM cosmology test, and bias in $w$ for the Flat $w$CDM test with strong prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$. This allows us to investigate biases without the investigative hindrances of non-Gaussian or truncated posterior surfaces. The strong prior on $\Omega_m$ cuts a slice through the traditional `banana' posterior surface in the $w$-$\Omega_m$ plane of Figure \ref{fig:simple_w}. Without making such a slice, the variation in $w$ is larger due to a shift along the degeneracy direction of the `banana'. By focusing the slice at an almost fixed $\Omega_m$, we can see the variation in the mean value of $w$ approximately perpendicular to the lines of degeneracy, instead of along them. The results of the analysis are detailed in Table \ref{tab:simple_model}, and demonstrate the performance of our model in recovering the true cosmological parameters.  As we are using 100 independent realisations, the precision of our determination of the mean simulation result is a tenth of the quoted scatter. The deviation from truth values is below this threshold, no significant bias is detected in either the Flat $\Lambda$CDM model or the Flat $w$CDM model. With this simple data, we also correctly recover underlying supernova populations, which can be seen in Figure \ref{fig:simple_w_super}.

@@

\begin{deluxetable}{lcccc}
\tablewidth{\columnwidth}

\tablecolumns{5}
\tablecaption{Supernova population distributions}
\tablehead {
	\colhead {Model} &
	\colhead { $\langle x_1 \rangle$  } &
	\colhead { $\sigma_{x_1}$ } &
	\colhead { $\langle c \rangle$ } &
	\colhead { $\sigma_c$ } 
}
	\startdata
		{\sk} low-$z$    &  $0.55$ \& $-1.5$ & $^{+0.45}_{-1.0}$ \& $^{+0.5}_{-0.5}$ & $-0.055$ & $^{+0.15}_{-0.023}$ \\ \\
		{\sk} DES     & $0.973$ & $^{+0.222}_{1.472}$ & $-0.054$  &  $^{+0.101}_{0.043}$  \\
	\enddata
	\label{tab:dist}
	\tablecomments{The {\sk} low-$z$ stretch distribution is formed as sum of two bifurcated Gaussians, with the mean and spread of each component given respectively.}
\end{deluxetable}

@@

\begin{deluxetable*}{lrrrrr}
	\tablewidth{\textwidth}
	\tablecolumns{6}
	\tablecaption{Realistic simulation standardisation parameters}
	\tablecomments{Standardisation parameters and base intrinsic scatter parameter results for the 100 fits to {\gten} and {\celeven} simulations. We show the average parameter mean and average standard deviation respectively, with the simulation scatter shown in brackets, such that each cell shows $\langle\mu\rangle\ [\langle\sigma\rangle\ ({\rm scatter})]$. The width of the intrinsic scatter ($\sigma_{m_B}$) does not have an input truth value as it is determined from the scatter model.}
	\tablehead{
		Model & $\alpha - \alpha_{\rm True}$ & $\beta - \beta_{\rm True}$ & $\langle M_B \rangle - \langle M_B \rangle_{\rm True}$ & $\sigma_{m_B}^{\rm DES}$ & $\sigma_{m_B}^{{\rm low-}z}$	
	}
	\startdata
		{\gten} Stat + Syst  & $0.022 \ [0.009\ (0.008)]$ & $ 0.34 \ [0.19\ (0.18)]$ & $-0.002 \ [0.028\ (0.015)]$ & $0.070 \ [0.022\ (0.018)]$ & $0.073 \ [0.025\ (0.022)]$\\
		{\gten}  Stat         & $0.000 \ [0.008\ (0.008)]$ & $0.33 \ [0.20\ (0.17)]$ & $0.001 \ [0.016\ (0.013)]$ & $0.069 \ [0.023\ (0.019)]$ & $0.072 \ [0.026\ (0.023)]$\\
		{\celeven} Stat + Syst  &  $0.002 \ [0.009\ (0.007)]$ & $-0.04 \ [0.15\ (0.13)]$ & $0.014 \ [0.030\ (0.018)]$ & $0.024 \ [0.016\ (0.011)]$ & $0.029 \ [0.020\ (0.014)]$ \\
		{\celeven} Stat         &  $0.000 \ [0.008\ (0.007)]$ & $-0.05 \ [0.16\ (0.13)]$ & $0.006 \ [0.016\ (0.015)]$ & $0.025 \ [0.016\ (0.012)]$ & $0.027 \ [0.020\ (0.015)]$ \\

	\enddata
	\label{tab:standardisation}
\end{deluxetable*}

@@

\begin{deluxetable}{lrr}
	\tablecaption{Realistic simulation determination of $w$}
	\tablewidth{\columnwidth}
	\tablecomments{Investigating the combined 100 fits to {\gten} and {\celeven} simulations, fitting with both statistics only and also when including systematics. The quoted value for $w$ represents the average mean of the fits, with the average uncertainty being shown in square brackets and the simulation scatter (the standard deviation of the mean of 100 fits) shown in standard brackets. The bias significance represents our confidence that the deviation in the mean $w$ away from $-1$ is not due to statistical fluctuation.}
	\tablehead{		Model & $w\ \langle\mu\rangle\ [\langle\sigma_w\rangle$ (scatter)$]$ & $w$-Bias		}
	\startdata
		{\gten} Stat + Syst  &  $-0.998 \ [0.097\ (0.073)]$ & $(0.02\pm0.07)\sigma$  \\
		{\gten}  Stat         &  $-1.008\ [0.080\ (0.068)]$ & $(-0.10\pm0.08)\sigma$   \\
		{\celeven} Stat + Syst  &  $-0.945\ [0.098\ (0.077)]$ & $(0.55\pm0.08)\sigma$  \\
		{\celeven} Stat         &  $-0.948\ [0.079\ (0.066)]$ & $(0.65\pm0.08)\sigma$   \\
	\enddata
		\label{tab:bulk_summary}
\end{deluxetable}

\begin{deluxetable}{crr}

	\tablecaption{Reduced parameter correlations with $w$.}
	\tablecomments{Correlations determined from the combined 100 simulation fits. Correlations for the low-$z$ band systematics and the latent parameters representing selection function uncertainty are not shown but have negligible correlation. Zero superscripts indicate the Dark Energy Survey, and a superscript one represents the low-$z$ survey.}
	\tablehead{		Parameter & {\gten} Stat+Syst & {\celeven} Stat+Syst }
	\startdata
		$\Omega_m$ &   $-$0.19 &     $-$0.21     \\
		$\alpha$ &   $-$0.17 &     $-$0.20     \\
		$\beta$ &   $-$0.29 &     $-$0.23     \\
		$\langle M_B \rangle$ &    0.68 &      0.66     \\
		$\delta(0)$ &    0.00 &      0.00     \\
		$\delta(\infty)/\delta(0)$ &    0.00 &      0.00     \\
		\\
		$\sigma_{\rm m_B}^{0}$ &    0.04 &      0.07     \\
		$\sigma_{\rm m_B}^{1}$ &    0.23 &      0.18     \\
		$\sigma_{x_1}^{0}$ &    0.04 &      0.03     \\
		$\sigma_{x_1}^{1}$ &    0.05 &      0.01     \\
		$\sigma_{c}^{0}$ &    0.01 &      0.11     \\
		$\sigma_{c}^{1}$ &    0.08 &      0.04     \\
		$\alpha_c^{0}$ &   $-$0.04 &      0.04     \\
		$\alpha_c^{1}$ &    0.03 &      0.01     \\
		$\kappa_{c0}^{0}$ &   $-$0.10 &     $-$0.05     \\
		$\kappa_{c0}^{1}$ &   $-$0.20 &     $-$0.17     \\
		$\kappa_{c1}^{0}$ &   $-$0.05 &     $-$0.01     \\
		$\kappa_{c1}^{1}$ &   $-$0.01 &      0.01     \\
		$\langle x_1^{0} \rangle$ &   $-$0.01 &     $-$0.05     \\
		$\langle x_1^{1} \rangle$ &   $-$0.02 &      0.02     \\
		$\langle x_1^{2} \rangle$ &   $-$0.04 &     $-$0.04     \\
		$\langle x_1^{3} \rangle$ &   $-$0.03 &     $-$0.06     \\
		$\langle x_1^{4} \rangle$ &   $-$0.06 &     $-$0.06     \\
		$\langle x_1^{5} \rangle$ &    0.04 &      0.02     \\
		$\langle x_1^{6} \rangle$ &    0.04 &      0.04     \\
		$\langle x_1^{7} \rangle$ &    0.08 &      0.03     \\
		$\langle c^{0} \rangle$ &   $-$0.05 &     $-$0.12     \\
		$\langle c^{1} \rangle$ &    0.11 &      0.03     \\
		$\langle c^{2} \rangle$ &    0.11 &      0.06     \\
		$\langle c^{3} \rangle$ &    0.14 &      0.04     \\
		$\langle c^{4} \rangle$ &   $-$0.11 &     $-$0.11     \\
		$\langle c^{5} \rangle$ &   $-$0.15 &     $-$0.08     \\
		$\langle c^{6} \rangle$ &   $-$0.12 &     $-$0.13     \\
		$\langle c^{7} \rangle$ &   $-$0.12 &     $-$0.06     \\
		\\
		$\delta [ {\rm SALT}_0 ]$ &    0.05 &      0.05     \\
		$\delta [ {\rm SALT}_1 ]$ &   $-$0.01 &      0.02     \\
		$\delta [ {\rm SALT}_2 ]$ &   $-$0.10 &     $-$0.09     \\
		$\delta [ {\rm SALT}_3 ]$ &   $-$0.03 &     $-$0.03     \\
		$\delta [ {\rm SALT}_4 ]$ &    0.08 &      0.09     \\
		$\delta [ {\rm SALT}_5 ]$ &    0.01 &      0.02     \\
		$\delta [ {\rm SALT}_6 ]$ &    0.05 &      0.07     \\
		$\delta [ {\rm SALT}_7 ]$ &   $-$0.11 &     $-$0.10     \\
		$\delta [ {\rm SALT}_8 ]$ &    0.01 &      0.02     \\
		$\delta [ {\rm SALT}_9 ]$ &    0.02 &      0.02     \\
		$\delta [ {\rm MWE}_{B-V} ]$ &    0.03 &      0.02     \\
		$\delta [ {\rm HST Calib} ]$ &   $-$0.07 &     $-$0.07     \\
		$\delta [ v_{\rm pec} ]$ &    0.00 &     $-$0.01     \\
		$\delta [ {\rm \delta z } ]$ &    0.01 &      0.00     \\
		$\delta [ \Delta g ]$ &    0.05 &      0.11     \\
		$\delta [ \Delta r ]$ &    0.16 &      0.10     \\
		$\delta [ \Delta i ]$ &   $-$0.16 &     $-$0.18     \\
		$\delta [ \Delta z ]$ &   $-$0.26 &     $-$0.26     \\
		$\delta [ \Delta \lambda_g ]$ &    0.16 &      0.20     \\
		$\delta [ \Delta \lambda_r ]$ &    0.05 &      0.06     \\
		$\delta [ \Delta \lambda_i ]$ &    0.00 &     $-$0.01     \\
		$\delta [ \Delta \lambda_z ]$ &    0.09 &      0.07     \\
	\enddata
	\label{tab:parameter_correlations}
\end{deluxetable}
